---
title: "R Notebook"
output: html_notebook
---

Computing arithmetic mean
```{r}
weather <- c(22, 24, 21, 22, 30, 28, 29)
mean(weather)
```


Computing weighted arithmetic mean
```{r}
weighted.mean(c(22.5, 27.7, 32.5), c(12/31, 18/31, 1/31))
```

Computing median
```{r}
median(weather)
```

Computing quantiles
```{r}
quantile(weather)
```

```{r}
quantile(weather, probs=c(0, 0.2, 0.4, 0.6, 0.8, 1))
```


Quantile-Quantile Plots (QQ-Plots)
```{r}
pizza <- read.csv('datasets/pizza_delivery.csv')
mario <- pizza[pizza$driver == 'Mario', ]
luigi <- pizza[pizza$driver == 'Luigi', ]
salvatore <- pizza[pizza$drive == 'Salvatore', ]
```

```{r}
qqplot(mario$time, salvatore$time)
abline(0, 1)
```

```{r}
qqplot(luigi$time, mario$time)
abline(0, 1)
```



(a) If all the pairs of quantiles lie (nearly) on a straight line at an angle of 45% from the x-axis, then the two samples have similar distributions.
(b) If the y-quantiles are lower than the x-quantiles, then the y-values have a tendency to be lower than the x-values.
(c) If the x-quantiles are lower than the y-quantiles, then the x-values have a tendency to be lower than the y-values.
(d) If there is a point where the x-quantiles are lower (or higher) than the y-quantiles for a specific range of values (a segment) it indicates that there is a break point up to which the y-quantiles are lower(or higher) than the x-quantiles and after that point, the y-quantiles are higher(or lower) than the x-quantiles.