---
title: "Association of two variables"
output: html_notebook
---

Creating a dummie dataset
```{r}
X = sample(x=20:35, size=100, replace=TRUE)
Y = sample(x=30:40, size=100, replace=TRUE)
```


Contingency table
```{r}
addmargins(table(X, Y))
```


Contingency table (relative frequencies)
```{r}
addmargins(prop.table(table(X,Y)))
```


# Bar charts
```{r}
air_class <- c(rep('1: Economy', 62), rep('2: Business', 25), rep('3: First', 13))
air_rating <- c(rep('1=poor', 10), rep('2=fair', 36), rep('3=Good', 40), rep('4=Very good', 14))

library(lattice)
barchart(table(air_class, air_rating), horizontal=FALSE, stack=FALSE)
barchart(table(air_class, air_rating), horizontal=FALSE, stack=FALSE)
```


To calculate the expected absolute frequencies in R, we can access the
"expected" object returned from a $\chi^2$-test applied to the respective contingency
table as follows:
```{r}
air_chisq <- chisq.test(table(air_class, air_rating))
air_chisq
```

```{r}
air_chisq$expected
```


Value of $\chi^2$ is between $0 \leq \chi^2 \leq n(\min(k, l) - 1)$
```{r}
air_chisq$statistic
```


Computing the Cramer's $V$ Statistic (Normalized $\chi^2$)
```{r}
library(vcd)
assocstats(xtabs(~ air_class + air_rating))
```

Computing the the (corrected version) Contingency Coefficient $C$
```{r}
library(vcd)
C_max = sqrt(
    (min(c(3, 4)) - 1) / min(c(3, 4))
)
assocstats(xtabs(~ air_class + air_rating))$cont / C_max
```


# Association Between Ordinal and Continuous Variables

```{r}
X <- c(10.85, 10.44, 10.50, 10.89, 10.62)
Y <- c(7.84, 7.96, 7.81, 7.47, 7.74)
```

## Pearson's Correlation Coefficient
```{r}
cor(X, Y, method='pearson')
```

## Spearman's Rank Correlation Coefficient
```{r}
cor(X, Y, method='spearman')
```


## Goodman and Kruskal's $\gamma$
```{r}
library(ryouready)
ex <- matrix(
    c(7, 11, 26, 10, 15, 31),
    ncol=3,
    byrow=T
)
ord.gamma(ex)
ord.tau(ex)
```


## Visualization  of Variables from Different Scales
Consider again our pizza delivery example. If we are interested in the pizza delivery times by branch, we may simply plot the box plots and ECDF’s of delivery time by branch. Next figure shows that the shortest delivery times can be observed in the branch in the East. Producing these graphs in R is straightforward: The boxplot command can be used for two variables by separating them with the ∼ sign. For the ECDF, we have to produce a plot for each branch and overlay them with the “add=TRUE” option.
```{r}
pizza <- read.csv('datasets/pizza_delivery.csv')
```

```{r}
boxplot(pizza$time ~ pizza$branch)
```

```{r}
plot.ecdf(pizza$time[pizza$branch == 'East'])
plot.ecdf(pizza$time[pizza$branch=='West'], add=TRUE)
plot.ecdf(pizza$time[pizza$branch=='Centre'], add=TRUE)
```



Exercise 4.8 Make yourself familiar with the Olympic decathlon data described in
Appendix A.4. Read in and attach the data in R.

(a) Use R to calculate and interpret the Bravais–Pearson correlation coefficient between the results of the discus and the shot-put events.
(b) There are 10 continuous variables. How many different correlation coefficients can you calculate? How would you summarize them?
(c) Apply the cor command to the whole data and interpret the output.
(d) Omit the two rows which contain missing data and interpret the output again.

```{r}
decathlon <- read.csv("datasets/decathlon.csv", row.names=1)
attach(decathlon)
cor(X.Discus,X.High.jump)
```

```{r}
cor(decathlon)
```

```{r}
cor(na.omit(decathlon))
```


```{r}
library(corrplot) # nice summary of results with package corrplot(not mentioned in solutions)
corrplot(cor(na.omit(decathlon)),method="number",col="black",cl.pos="n",tl.col="black")

detach(decathlon)
```



Exercise 4.9 We are interested in the pizza delivery data which is described in Appendix A.4.
(a) Read in the data and create two new binary variables which describe whether a pizza was hot (>65◦C) and the delivery time was short (<30min). Create a contingency table for the two new variables.
(b) Calculate and interpret the odds ratio for the contingency table from (a).
(c) UseCramer’sV,Stuart’sτc,GoodmanandKruskal’sγ,andastackedbarchartto explore the association between the categorical time and temperature variables.
(d) Drawascatterplotforthecontinuoustimeandtemperaturevariables.Determine
both the Bravais–Pearson and Spearman correlation coefficients.
(e) Use methods of your choice to explore the relationship between temperature and driver, operator, number of ordered pizzas and bill. Is it clear which of the
variables influence the pizza temperature?

```{r}
pizza <- read.csv("datasets/pizza_delivery.csv")
pizza$tempcat <- cut(pizza$temperature, breaks=c(0,65,100))
pizza$timecat <- cut(pizza$time, breaks=c(0,30,100))
attach(pizza)
```

a)
```{r}
addmargins(table(tempcat,timecat))
```

b)
```{r}
(101*261)/(213*691)
```

c)
```{r}
library(vcd)          
library(ryouready) 
library(lattice)      
assocstats(xtabs(~tempcat+timecat))
ord.gamma(table(tempcat,timecat))
ord.tau(table(tempcat,timecat))
```


d)
```{r}
plot(time,temperature)
```

e)
```{r}
boxplot(temperature~driver)
boxplot(temperature~operator)
```
